\chapter*{Acknowledgements}

\addcontentsline{toc}{chapter}{Acknowledgements}
\mtcaddchapter  

And so, as every journey, this one comes to an end too. 
A journey started 3 years ago, a morning of january 2017: it is incredible that 3 years have already been passed since then. 
During this time a lot of people supported me: tenth of pages will not be enough to mention all, but I am in duty to thanks some of them in these short lines. 

First of all, special mention to my Ph.D. directors and supervisors: Nathalie Bartoli, Peter Schmollgruber, Emmanuel Benard and Joseph Morlier. 
Without you, this achievement would have never seen an end. 
I hope that you had a great time as I did working with you. 

Thanks to prof. Dimitri Mavris and dr. Trevor Young, who reviewed the work, and to the jury members, who joined the Ph.D. defense: prof. Yves Gourinat, prof. Christian Bes, dr. Roelof Vos, and prof. Fabrizio Nicolosi. 

Thanks to all the colleagues from ONERA, who take part to the Ph.D. with their suggestions and advice: Sebastien Defoort, Michel Ridel, David Donjat, Remi Lafage, Julie Gauvrit-Ledogar, Sylvain Dubreuil. 

Thanks to Serge Bonnet, Marylene Dufort and Guillame Renaud from AIRBUS, for their collaboration in the Chair CEDAR framework. 

Thanks also to prof. Martins and John Jasa for the MDO Lab. at University of Michigan, as well as Justin Gray from NASA Glenn Research Centre and John Hwang from University of California San Diego, for their contribution to the Ph.D., and a special mention to the Formaction Doctorale of ISAE-Supaero who funded the visit at MDO Lab. in 2018. 

My neverending gratitude to the MAE students from ISAE-Supaero I had the honour to supervise: Larissa Cristina e Cunha Lima, Diana Aboub Karhim, Luca Cerquetani, Pablo Rodriguez Otiero, Harleen Kaur, Paolo Traverso, Sesha Chandini. 
Without the work you did in the Blended Wing-Body project, this reseach would have never seen the light. 

And now, let pass to some more personal acknowledgements. 
At the beginning, my intention was to simply write ``to me'', meaning that this achievement was due to myself, but none is so far from reality. 
If I arrived at this point, it is mainly because of my friend, who have always been from my side. 
First, thanks to all the people from ONERA: Morgane Menz, Gabriel Sarazin, Remy Priem, Marie Chiron, Joan Mas Colomer, Julien Marine, Lynda Ait Oubelli. 
You were more than colleagues, you were friends.

To Alessia, who did not hesitate to take a plane to be here with me in this day, you are really special, and you do not stop to show me that, day by day. 

To Nicola and Alex: our friendship last since our childhood, and despite the kilometers that divide us they will never end to stay close with me. 

To my beloved Antonella, one of my favourite people in the world, as a source of unlimited suggestions and comfort.

Last, but absolutely not least, to Sara, who is the demonstration that sometimes sisterhood is a matter of heart and not blood: thank you for have never given up on me, and trusted me even when I did not. 
This is for you. 

Finally, to my parents and my brother: the family is always a sure nest, and they always gave me all the support possible. 

Well now, was that all of them? 
Only one thing is certain, the best is yet to come.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../phdthesis"
%%% End: 
